Tail Indices and Scale Parameters in Financial Time Series

### 2017-02-28 papers/FX/draft.tex
Still struggling about K.
Started to correct the mistake related to the probability.

\begin{abstract}
We consider an investor with preferences that accord with Generalized
Disappointment Aversion. Such an investor cares about downside risk and
we assume he recognizes the heavy tail feature of asset return
distributions. We argue that when a market is dominated by rational
investors of this kind, the return distributions of equities that are
actively traded in this market must have very similar tail indices.

@@

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x < 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation
\eqref{eq:xxie0} can be re-arranged as
\[
u(v) + b u(\delta v))
=
\E u(C) + b \E [u(C); C < \delta v]
\]
The right-hand-side of the above equation is a functional of the
distribution function of $X$ via eq.\eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v]
\]
$C < \delta v$ implies
\[
X < -q(r, \phi, \delta v) := \ln \left( {
  \delta v - (1 - \phi) e^r
  \over
  \phi
} \right)
\]
In the rest of this section we write $q$ for $q(r, \phi, \delta v)$
for convenience. We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  -{\alpha K^\alpha \over \xi} p
  \int_{-\infty}^0
      {
       1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
      }
  {1 + b \1{x < -q} \over (K - x)^{\alpha + 1}} dx \nonumber \\
  && -{\beta K'^\beta \over \xi} (1 - p)
  \int_{0}^\infty
  {
   1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
  }
  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx
  \label{eq:xxie1.0}
\end{eqnarray}

@@

@@

  \item When $q \geq 0$, we have

@@

      && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\

@@

  \item When $q < 0$, it follows from \eqref{eq:xxie1.0}:
    \begin{eqnarray*}
      && \opd{\alpha}\mathcal U(F)(\alpha, \beta, K, K', \hat\phi) \\
      &=& \opd{\alpha} \left[
        -{\alpha K^\alpha \over \xi} p
        \int_{-\infty}^0
            {
              1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
            }
            {1 + b \over (K + x)^{\alpha + 1}} dx
        \right]
    \end{eqnarray*}
  \end{enumerate}
  So the same calculation as in the previous case carries through
  with 0 substituted for $q$. The same conclusion of
  \[
  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
  \]
  stands.
\end{proof}
Since $u(v) + b u(\delta v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$, all agents guided by this class of utility
functions will seek the largest $\alpha$ found in a stock group e.g.
a sector. As a result, one can expect all the actively traded equities
in the group to feature the same $\alpha$. In section
\ref{sec:data_analysis} we analyse real-world data of stock returns to
verify this conclusion.

In comparison to the theoretical constancy of $\alpha$, the scale and
shift parameter $K$ in the lower-tail distribution of $X$ may vary
hugely from one equity to another. This could be understood also by
computing $d\mathcal U F / dK$:

>>>>

\begin{abstract}
We consider an investor with preferences that accord with Generalized
Disappointment Aversion. Such an investor cares about downside risk and
we assume he recognizes the heavy tail feature of asset return
distributions. We argue that when a market is dominated by rational
investors of this kind, the return distributions of equities that are
actively traded in this market must have very similar tail indices. We
also show, in contrast, the scale parameters of the return
distributions may differ hugely from one another.

@@

\subsubsection[in case gamma > 1]{in case $\gamma > 1$}
In this section we consider power utility functions of
\eqref{eq:power_utility} with $\gamma > 1$. For convenience,
let $\xi = \gamma - 1$. In this case, we assume the following
distribution function $F(\cdot)$ of the equity return:
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
Equation
\eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(-q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q(r, \phi) &:=& -\ln \left( {
    \delta v - (1 - \phi) e^r
    \over
    \phi
} \right)
\end{eqnarray*}
Note $C < \delta v$ implies
\[
X < -q(r, \phi)
\]
The right-hand-side of \eqref{eq:functional} is a functional of the
distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(-q)
\]
In the rest of this section we write $q$ for $q(r, \phi)$
for convenience. We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=& \underbrace{
  -{\alpha K^\alpha \over \xi} p
  \int_{-\infty}^0
      {
       1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
      }
  {1 + b \1{x < -q} \over (K - x)^{\alpha + 1}} dx
  }_{J_1} \nonumber \\
  && -{\beta K'^\beta \over \xi} (1 - p)
  \int_{0}^\infty
  {
   1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
  }
  {1 + b \1{x < -q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  &&
  \underbrace{
    -b u(\delta v) p {K^\alpha \over (K + q)^\alpha} \1{q \geq 0}
  }_{J_2}
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      \over
      (K' - q)^\beta
    } \1{q < 0}
  \right]
  \label{eq:xxie1.0}
\end{eqnarray}

@@

    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &=& \pd{J_1}{\alpha} + \pd{J_2}{\alpha} > 0
  The 2nd term is easily computed:
    \pd{J_2}{\alpha} &=& b u(\delta v) p \1{q \geq 0}
      {K \over K + q}
    \right)^\alpha
      1 + {q \over K}
    \right) > 0
  As for the 1st term we have

@@

  \item When $q \geq 0$,

@@

      && \pd{J_1}{\alpha}(\alpha, \beta, K, K', \hat \phi) \\

@@

  \item When $q < 0$, it follows from \eqref{eq:xxie1.0}:
    \begin{eqnarray*}
      && \opd{\alpha}J_1(\alpha, \beta, K, K', \hat\phi) \\
      &=& \opd{\alpha} \left[
        -{\alpha K^\alpha \over \xi} p
        \int_{-\infty}^0
            {
              1 \over \left[ (1 - \phi) e^r + \phi e^x \right]^{\xi}
            }
            {1 + b \over (K + x)^{\alpha + 1}} dx
        \right]
    \end{eqnarray*}
  \end{enumerate}
  So the same calculation as in the previous case carries through
  with 0 substituted for $q$. The same conclusion of
  \[
  \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) > 0
  \]
  stands.
\end{proof}
Since $u(v) + b u(\delta v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$, all agents guided by this class of utility
functions will seek the largest $\alpha$ found in a stock group e.g.
a sector. As a result, one can expect all the actively traded equities
in the group to feature the same $\alpha$. In section
\ref{sec:data_analysis} we analyse real-world data of stock returns to
verify this conclusion.

In comparison to the theoretical constancy of $\alpha$, the scale and
shift parameter $K$ in the lower-tail distribution of $X$ may vary
hugely from one equity to another. This could be understood by observing
\[
\lim_{K \to 0} F(x) = \left\{
\begin{array}{ll}
  p \delta_0((-\infty, x])) & \text{ if } x \leq 0 \\
  1 - (1 - p) \left(
    K' \over K' + x
  \right)^\beta & \text{ if } x > 0
\end{array}
\right.
\]
